Final Model Analysis

### 2017-08-27 model fitting/final_model.Rmd
plotted predicted survival estimates

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
x <- oshitar::setup()
library(survival); library(dplyr); library(rms)
```

```{r}
dir <- file.path(getwd(),"data")
out <- read.csv(system.file("extdata/answers_data.csv", package = "oshitar"))

finaldata <- oshitar::variable_setup(out)
identical(finaldata$contain_answered, x$contain_answered)

write.csv(finaldata, file = "final_data.csv")
```

@@

```{r, echo = FALSE}
# building model on train1
trainmodel <- cph(Surv(time_until_answer, answered) ~ new_category + new_user + contain_unanswered + contain_answered + title_questionmark + title_beginwh + text_contain_punct + text_all_lower + update + greeting + gratitude + prior_effort + weekday + strat(ampm) + sqrt(avg_tag_score) + poly(text_length, 2) + rcs(device_length, 5) + rcs(avg_tag_length, 4) + rcs(newline_ratio, 4), data = train1, x = TRUE, y = TRUE)
AIC(trainmodel, k = 2) # 54428.62
trainmodel
```

@@

@@

```{r, echo = FALSE}
vars <- "new_category + new_user + contain_unanswered + contain_answered + title_questionmark + title_beginwh + text_contain_punct + text_all_lower + update + greeting + gratitude + prior_effort + weekday + strat(ampm) + sqrt(avg_tag_score) + poly(text_length, 2) + rcs(device_length, 5) + rcs(avg_tag_length, 4) + rcs(newline_ratio, 4)"

@@

###### output: 
* Dxy = Somers' Dxy (2 ∗ (C − 0.5), where C is the concordance probability)
* R2: RN2 index
* Slope: Calibration slope (slope of predicted log odds vs true log odds) optimism value is a value of overfitting
* D: Discrimination index — likelihood ratio χ2 divided by the sample size
* U: Unreliability index — unitless index of how far the logit calibration curve intercept and slope are from (0, 1)
* Q: Logarithmic accuracy score - scaled version of the log-likelihood achieved by the predictive model
* g: g index (?)

###### interpreting:
* optimism is small- model is a good fit (?)
* slope shrinkage = 0.0163, indicates over-fitting not present

```{r, echo = FALSE}
valmodel <- rms::cph(Surv(time_until_answer, answered) ~ new_category + new_user + contain_unanswered + contain_answered + title_questionmark + title_beginwh + text_contain_punct + text_all_lower + update + greeting + gratitude + prior_effort + weekday + strat(ampm) + sqrt(avg_tag_score) + poly(text_length, 2) + rcs(device_length, 5) + rcs(avg_tag_length, 4) + rcs(newline_ratio, 4), data = x, x = TRUE, y = TRUE, surv = TRUE, time.inc = 0.5, dxy = TRUE)

validation <- rms::validate(valmodel, method = "crossvalidation", B = 100, bw = F, dxy = TRUE, u = 0.5)
validation
```

@@

```{r, echo = FALSE}
# calibration accuracy in predicting probability at 0.5 hours (predicts and calibrates at 0.5 hours for each iteration of the bootstrap?)
calibration <- rms::calibrate(calmodel, method = "crossvalidation", B = 100, bw = FALSE, u = 0.5)
calibration
plot(calibration)
```

@@

```{r, echo = FALSE}
model <- rms::cph(Surv(time_until_answer, answered) ~ new_category + new_user + contain_unanswered + contain_answered + title_questionmark + title_beginwh + text_contain_punct + text_all_lower + update + greeting + gratitude + prior_effort + weekday + strat(ampm) + sqrt(avg_tag_score) + poly(text_length, 2) + rcs(device_length, 5) + rcs(avg_tag_length, 4) + rcs(newline_ratio, 4), data = x, x = TRUE, y = TRUE, surv = TRUE)

x$predictions <- exp(predict(model, type = "lp"))

@@

```{r, echo = FALSE}
surv_prob <- model$surv # list of 4 (one list for each stratification level)

# allows you to get estimated survival probability at specific time (for each strat level)
summary(survfit(model), time = c(0.5, 1, 3))

# either specify newdata, linear.predictors or x
# number on the rows of the surv matrix is the index 
# function is very slow if full data set is used
rms::survest(model, newdata = test1[1:3,], times = c(0.5, 1, 10), conf.int = FALSE)

# using different function
pec::predictSurvProb(model, newdata = test1[1:3,], times = c(0.5, 1, 10))

times <- sort(unique(x$time_until_answer))

pec::plotPredictSurvProb(model, newdata = x[1:4,], times = times) # this is ugly, figure out how to do this in ggplot?
```

##### Describing the final model

* predicted survival curve + curves for certain variables (rms functions to do this keep outputting errors about datadist)
* nomograms?

@@

```{r}
# input vector of times to predict on + data set to be used in predictions
# function runs really slow
# outputs matrix of predicted survival probabilities 
# also include some kind of performance metric? measure of how accurately its predicting 
# sets up variables needed in the new data 
get_survival <- function(times, data) {
  # setting up necessary variables in new data 
  newdata <- oshitar::variable_setup(data)

@@

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
x <- read.csv("final_data.csv", head = TRUE)
library(survival); library(dplyr); library(rms)
```

@@

```{r, echo = FALSE}
# building model on train1 (df set to 5 for all splines at first, but setting tag and newline to 4 minimized AIC and performed best)
trainmodel <- cph(Surv(time_until_answer, answered) ~ new_category + new_user + contain_unanswered + contain_answered + title_questionmark + title_beginwh + text_contain_punct + text_all_lower + update + greeting + gratitude + prior_effort + weekday + strat(ampm) + sqrt(avg_tag_score) + pol(text_length, 2) + rcs(device_length, 5) + rcs(avg_tag_length, 4) + rcs(newline_ratio, 4), data = train1, x = TRUE, y = TRUE)
AIC(trainmodel, k = 2) # 54421.53
trainmodel
```

@@

  # computing performance metrics 
  metric1 <- rms::cph(Surv(time_until_answer, answered) ~ predictions, data = test)
  test_metrics <- data.frame(HR = exp(metric1$coefficients), 
                             LR = round(metric1$stats[3],2), 
                             pval = round(metric1$stats[5],2), 
                             R2 = round(metric$stats[8], 2),
                             AIC = stats::AIC(metric1, k = 2),
                             Dxy = round(metric1$stats[9],2), 
                             Concordance = survConcordance(Surv(time_until_answer, answered) ~ predictions, data = test)$concordance)

@@

```{r, echo = FALSE}
vars <- "new_category + new_user + contain_unanswered + contain_answered + title_questionmark + title_beginwh + text_contain_punct + text_all_lower + update + greeting + gratitude + prior_effort + weekday + strat(ampm) + sqrt(avg_tag_score) + pol(text_length, 2) + rcs(device_length, 5) + rcs(avg_tag_length, 4) + rcs(newline_ratio, 4)"

@@

###### output:

* Dxy = Somers' Dxy (2 ∗ (C − 0.5), where C is the concordance probability)
* R2: RN2 index
* Slope: Calibration slope (slope of predicted log odds vs true log odds) optimism value is a value of overfitting
* D: Discrimination index — likelihood ratio χ2 divided by the sample size
* U: Unreliability index — unitless index of how far the logit calibration curve intercept and slope are from (0, 1)
* Q: Logarithmic accuracy score - scaled version of the log-likelihood achieved by the predictive model
* g: g index (?)

###### interpreting:

* optimism is small- model is a good fit (?)
* slope shrinkage = 0.0163, indicates over-fitting not present

```{r, echo = FALSE}
dd = datadist(x)
options(datadist = "dd")

final <- rms::cph(Surv(time_until_answer, answered) ~ new_category + new_user + contain_unanswered + contain_answered + title_questionmark + title_beginwh + text_contain_punct + text_all_lower + update + greeting + gratitude + prior_effort + weekday + strat(ampm) + sqrt(avg_tag_score) + pol(text_length, 2) + rcs(device_length, 5) + rcs(avg_tag_length, 4) + rcs(newline_ratio, 4), data = x, x = TRUE, y = TRUE, surv = TRUE, time.inc = 0.5, dxy = TRUE)

validation <- rms::validate(final, method = "crossvalidation", B = 100, bw = F, dxy = TRUE, u = 0.5)
validation
```

@@

```{r, echo = FALSE}
# calibration accuracy in predicting probability at 0.5 hours (predicts and calibrates at 0.5 hours for each iteration of the bootstrap?)
calibration <- rms::calibrate(final, method = "crossvalidation", B = 100, bw = FALSE, u = 0.5)
calibration
plot(calibration)
```

@@

```{r, echo = FALSE}
x$predictions <- exp(predict(final, type = "lp"))

@@

```{r, echo = FALSE}
surv_prob <- final$surv # list of 4 (one list for each stratification level)

# allows you to get estimated survival probability at specific time (for each strat level)
summary(survfit(final), time = c(0.5, 1, 3))

# either specify newdata, linear.predictors or x
# number on the rows of the surv matrix is the index 
# function is very slow if full data set is used
rms::survest(final, newdata = test1[1:3,], times = c(0.5, 1, 10), conf.int = FALSE)

# using different function
pec::predictSurvProb(final, newdata = test1[1:3,], times = c(0.5, 1, 10))
```

##### Estimated survival probability curves

* predicted survival curve + curves for certain variables (rms functions keep outputting errors about data dist)
* nomograms?

```{r}
# plots log hazard for each variable 
plots <- ggplot(Predict(final)) # output is messy, figure out how to access each plot separately

times <- sort(unique(x$time_until_answer))
pec::plotPredictSurvProb(final, newdata = x[1:4,], times = times) # ugly output

# must specify a variable 
rms::survplot(final, ampm, time.inc = 100, col = 1:4, n.risk = TRUE)

# plot the effects of predictors
plot(Predict(final, new_category))
plot(Predict(final, new_user))
plot(Predict(final, new_user, new_category))

@@

```{r}
# input vector of times to predict on + data set to be used in predictions
# function runs really slow
# outputs matrix of predicted survival probabilities 
# sets up variables needed in the new data 
get_survival <- function(times, data) {
  # setting up necessary variables in new data 
  newdata <- oshitar::variable_setup(data)

@@

get_survival(times = c(0.5, 1, 10), data = out[1:10,])
```
